Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-07-27 manuscript/usses_water_ms.Rmd
adds references

Second, ANPP at our site may be influenced by factors beyond the window of soil moisture we included in our statistical model.
For example, temperature can also impact community composition and ANPP [@Hoover2014].
Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
Likewise, we did not redistribute snow across our treatments in the winter, and snowmelt may spur early spring growth (CITATION).

Third, the life history traits of the dominant species in our study ecosystem may explain the consistent and weak effect of soil moisture on ANPP (Fig. 3).
Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success (CITATION).
The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the credible intervals for $\beta_1$ overlapping zero), even at precipitation extremes.

Another strategy to deal with variable environmental conditions is avoidance (CITATION), which would result in a consistent ecosystem functional response at low soil moisture.
The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R. Kleinhesselink, unpublished data].
For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system (CITATION).

>>>>

Second, ANPP at our site may be influenced by factors beyond the window of soil moisture we included in our statistical model.
For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
Likewise, we did not redistribute snow across our treatments in the winter, and snowmelt may spur early spring growth.
These statistical issues of missing potentially important covariates could explain the weak and uncertain relationship we observed between soil moisture and ANPP.

Third, the life history traits of the dominant species in our study ecosystem may explain the consistent but weak and uncertain effect of soil moisture on ANPP (Fig. 3).
Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success [@Seger1987].
In other words, species do the same thing every year, with only minimal response to environmental conditions.
The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the credible intervals for $\beta_1$ overlapping zero), even at precipitation extremes.

Another strategy to deal with variable environmental conditions is avoidance, which would result in a consistent ecosystem functional response at low soil moisture.
The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R. Kleinhesselink, unpublished data].
For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].

### 2017-07-27 manuscript/usses_water_ms.tex
adds references

Second, ANPP at our site may be influenced by factors beyond the window
of soil moisture we included in our statistical model. For example,
temperature can also impact community composition and ANPP
\citep{Hoover2014}. Measurements of soil moisture likely contain a
signal of temperature, through its impact on evaporation and
infiltration, but the measurements will not capture the direct effect of
temperature on metabolic and physiological processes. Likewise, we did
not redistribute snow across our treatments in the winter, and snowmelt
may spur early spring growth (CITATION).

Third, the life history traits of the dominant species in our study
ecosystem may explain the consistent and weak effect of soil moisture on
ANPP (Fig. 3). Species that live in variable environments, such as cold
deserts, must have strategies to ensure long-term success as conditions
vary. One strategy is bet hedging, where species forego short-term gains
to reduce the variance of long-term success (CITATION). The dry and
variable environment of the sagebrush steppe has likely selected for bet
hedging species that can maintain function at low water availability and
have weak responses to high water availability. In so doing, the
dominant species in our ecosystem avoid ``boom and bust'' cycles, which
corresponds to the weak effect of soil moisture on ANPP (i.e., the
credible intervals for \(\beta_1\) overlapping zero), even at
precipitation extremes.

Another strategy to deal with variable environmental conditions is
avoidance (CITATION), which would result in a consistent ecosystem
functional response at low soil moisture. The perennial plants in this
cold desert ecosystem are tolerant to drought conditions \citep[A.R.
Kleinhesselink, unpublished data]{Bazzaz1979, Franks2011}. For example,
many of the perennial grasses in our focal ecosystem avoid drought
stress by growing early in the growing season \citep[A.R.
Kleinhesselink, personal observation]{Blaisdell1958}. Likewise, the
dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has
access to water deep in the soil profile thanks to a deep root system
(CITATION).

>>>>

Second, ANPP at our site may be influenced by factors beyond the window
of soil moisture we included in our statistical model. For example,
temperature can impact ANPP directly \citep{Epstein1997} and by
exacerbating the effects of soil moisture \citep{DeBoeck2011}.
Measurements of soil moisture likely contain a signal of temperature,
through its impact on evaporation and infiltration, but the measurements
will not capture the direct effect of temperature on metabolic and
physiological processes. Likewise, we did not redistribute snow across
our treatments in the winter, and snowmelt may spur early spring growth.
These statistical issues of missing potentially important covariates
could explain the weak and uncertain relationship we observed between
soil moisture and ANPP.

Third, the life history traits of the dominant species in our study
ecosystem may explain the consistent but weak and uncertain effect of
soil moisture on ANPP (Fig. 3). Species that live in variable
environments, such as cold deserts, must have strategies to ensure
long-term success as conditions vary. One strategy is bet hedging, where
species forego short-term gains to reduce the variance of long-term
success \citep{Seger1987}. In other words, species do the same thing
every year, with only minimal response to environmental conditions. The
dry and variable environment of the sagebrush steppe has likely selected
for bet hedging species that can maintain function at low water
availability and have weak responses to high water availability. In so
doing, the dominant species in our ecosystem avoid ``boom and bust''
cycles, which corresponds to the weak effect of soil moisture on ANPP
(i.e., the credible intervals for \(\beta_1\) overlapping zero), even at
precipitation extremes.

Another strategy to deal with variable environmental conditions is
avoidance, which would result in a consistent ecosystem functional
response at low soil moisture. The perennial plants in this cold desert
ecosystem are tolerant to drought conditions \citep[A.R. Kleinhesselink,
unpublished data]{Bazzaz1979, Franks2011}. For example, many of the
perennial grasses in our focal ecosystem avoid drought stress by growing
early in the growing season \citep[A.R. Kleinhesselink, personal
observation]{Blaisdell1958}. Likewise, the dominant shrub in our focal
ecosystem, \emph{Artemisia tripartita}, has access to water deep in the
soil profile thanks to a deep root system \citep{Kulmatiski2017a}.
